---
title: "nn2-timings"
author: "Greg Jefferis"
output: html_document
---

## Introduction

This vignette gives some indication of how kd build and query times
vary with the length of the query, target input data. These are based
on real world data for Drosophila neurons 
(see [Costa et al 2016](https://doi.org/10.1016/j.neuron.2016.06.012)).
Since the data are large, they are not distributed with the package
but may be obtained on request.

```{r options, include=FALSE}
# only run locally - this vignette takes a long time and
# currently relies on some large (50 Mb) sample data.
knitr::opts_chunk$set(eval=nzchar(Sys.getenv("RUN_RANN2_VIGNETTES")))
```

The data used here 

```{r setup, message=FALSE}
library(knitr)
library(nat)
rgl::setupKnitr()
library(microbenchmark)
library(RANN2)
library(plyr)
dps=readRDS("~/projects/ChiangReanalysis/data/dpscanon.rds")
```

```{r sample}
set.seed(42)
s=sample(dps,500)
library(nat.nblast)
np=neuron_pairs(s, n = 500)
sl=sapply(s, function(x) nrow(x$points))
np$query.len=sl[np$query]
np$target.len=sl[np$target]
```

Let's investigate the build time for the kdtree of those neurons
```{r buildtime, cache=TRUE}
times=llply(np$target, function(n)
  microbenchmark(nn2(s[[n]]$points,matrix(0,ncol=3,nrow=1), k=1), times = 20))
```

```{r model-buildtimes, message=FALSE}
np$target.time=sapply(times,function(x) median(x$time)/1e6)
library(ggplot2)
qplot(target.len,target.time, data=np, ylab='Tree build time /ms') + 
  scale_x_log10()+scale_y_log10()+geom_smooth()
l=lm(target.time~target.len, data=np)
l$coefficients[2]
```

So in conclusion, ~ 0.5 ms / 1000 points to build.

For queries, total time will include build time, + query time depending O(n)
on number of query points and O(n log n) on number of target points.

```{r searchtime, cache=TRUE}
times2=mlply(np, function(query, target, ...)
  microbenchmark(nn2(s[[target]]$points,s[[query]]$points, k=1), times = 20))
np$search.time=sapply(times2,function(x) median(x$time)/1e6)
```

```{r plot-searchtime, fig.width=7, fig.height=6, webgl=TRUE}
with(np, plot3d(log2(query.len), log2(target.len), log2(search.time)))
```

```{r plot-searchtime-1-neuron}
set.seed(421)
s20=sample(s, 20)
times3=llply(s20, function(query, ...)
  microbenchmark(nn2(s20[[1]]$points,query$points, k=1), times = 20))
df=data.frame(query=names(s20),target= names(s20)[1], stringsAsFactors = F)
df$search.time=sapply(times3,function(x) median(x$time)/1e6)
df$query.len=sl[df$query]
qplot(query.len, search.time, data=df, ylab='Query time /ms') +
  geom_smooth(method = 'lm')
l=lm(search.time~query.len, data=df)
```

Slope is `r l$coefficients[2]`.

```{r plot-searchtime-1-neuron-target}
set.seed(421)
s20=sample(s, 20)
np40=neuron_pairs(s20[1:2],s20)
times4=mlply(np40, function(target, query, ...)
  microbenchmark(nn2(s20[[target]]$points, s20[[query]]$points, k=1), times = 20))

np40$search.time=sapply(times4,function(x) median(x$time)/1e6)
np40$target.len=sl[np40$target]
np40$query.len=sl[np40$query]
qplot(target.len, search.time, data=np40, ylab='Query time /ms') +
  geom_smooth(method = 'lm')
l=lm(search.time~target.len, data=np40)
```

Slope is `r l$coefficients[2]`.
